---
title: "Sex_Buy_SSVD"
author: "Sarah Van Dyk"
date: "2024-11-30"
output: html_document
---

#Part One Cleaning Data

##Load Library and Data 
```{r}
#Load Packages
library(tidyverse)
library(haven)
library(gt)


#Load Data
setwd("~/Library/CloudStorage/Box-Box/Sarah Sowell UG-HopeNet Datasets")
sexBuy <- read_dta("72mo_updated_20220816_SexBehavior-Violence_v0.dta")

head(sexBuy)
table(sexBuy$female)


#Set Aside Second DF for Data w/ Females
sexBuy_all <- sexBuy


#Remove Women
sexBuy<- subset(sexBuy, (sexBuy$female == 0))



```
#Summary Statistics
```{r}
#Create Variable TransSex (Main DV)
sexBuy$transSex <- ifelse(!is.na(sexBuy$pastyr_sexworker) & sexBuy$pastyr_sexworker == 1, 1,
                   ifelse(!is.na(sexBuy$pastyr_gavegiftsex) & sexBuy$pastyr_gavegiftsex == 1, 1,0))

#Create IVs needed to filter data

#Binge Drinking
sexBuy$binge_drink <- ifelse(sexBuy$whenalcohol > 3,1,0)
sexBuy$binge_drink <- ifelse(is.na(sexBuy$binge_drink), 0, sexBuy$binge_drink)


#Early Debut Binary 
sexBuy$early_sex_debut <- ifelse(sexBuy$firstsex_age < 15, 1,0)

#Forced Debut Binary
sexBuy$nonconsent_start <- ifelse(sexBuy$firstsex_voluntary ==1,0,1)

#Current Abuse# Compute 'current_abuse' as the sum of 'forcesex_nonphys' and 'forcesex', and add 1
sexBuy$current_abuse <- rowSums(sexBuy[c("forcesex_nonphys", "forcesex")], na.rm = TRUE) 

# Create 'current_abuse_b' based on the 'current_abuse' value
sexBuy$current_abuse_b <- ifelse(sexBuy$current_abuse > 0, 1, 0)


#Historical Abuse
sexBuy$historical_sex_abuse <- rowSums(sexBuy[c("early_sex_debut", "nonconsent_start", "ace_5", "ace_6")] >0, na.rm = TRUE)

# Create 'hist_sex_b' based on the value of 'historical_sex_abuse'
sexBuy$hist_sex_b <- ifelse(sexBuy$historical_sex_abuse > 0, 1, 0)


#Physical Violence
sexBuy$phys_violence_index <- rowSums(sexBuy[c("slap", "push", "kick", "weapon", "insult", "threaten", "twisted", "choke")], na.rm = TRUE)

# Create 'hphys_viol_b' based on the value of 'phys_violence_index'
sexBuy$phys_viol_b <- ifelse(sexBuy$phys_violence_index > 0, 1, 0)

#Create Abuse Variable
sexBuy$newAbuse <- rowSums(sexBuy[, c("ace_1", "ace_2", "ace_3", "ace_4")] > 0)

table(sexBuy$newAbuse)


#Subset only by men included in final analysis 
sexBuy <- sexBuy %>%
  dplyr::filter(!is.na(transSex), 
                !is.na(binge_drink), 
                !is.na(ace_aphys), 
                !is.na(current_abuse), 
                !is.na(historical_sex_abuse), 
                !is.na(phys_viol_b))



```

##Analysis Distribution of Background Charateristics 
```{r}
#Martial Status 
table(sexBuy$mstatus)
489/694

#Age
sexBuy$age_below_30 <- ifelse(sexBuy$age < 31, 1,0)
table(sexBuy$age_below_30)
493/694

#Sex Ever
table(sexBuy$sex_ever)
21/(21+673)

#Days Drinking 
sexBuy$alcdays_w_no_drink <- ifelse(is.na(sexBuy$alcoholdays), 0, sexBuy$alcoholdays)
table(sexBuy$alcdays_w_no_drink)
mean(sexBuy$alcdays_w_no_drink, na.rm=TRUE)
sd(sexBuy$alcdays_w_no_drink, na.rm=TRUE)
385/694


#HIV Status
table(sexBuy$hivstat)
67/694

#HIV Risk
table(sexBuy$hiv_risk)
71/(71+181+363)

names(sexBuy)
#Total Aces
table(sexBuy$total_ace)
74/694
13/694
mean(sexBuy$total_ace)
sd(sexBuy$total_ace)

#Sexual Abuse
table(sexBuy$hist_sex_b)
62/694

#Sexual Abuse
table(sexBuy$current_abuse_b)
58/694

table(sexBuy$hist_sex_b,sexBuy$current_abuse_b)
8/694
```

##Create Table of Summary Stats for Non-Binary Variables
```{r}
#Create Needed Variables
sexBuy$married <- ifelse(sexBuy$mstatus >1,0,1)


# Calculate overall averages for alcoholdays, hiv_risk, and total_ace
overall_stats <- tibble(
  avg_alcoholdays = round(mean(sexBuy$alcdays_w_no_drink, na.rm = TRUE), 1),
  sd_alcoholdays = round(sd(sexBuy$alcoholdays, na.rm = TRUE), 1),
  avg_hiv_risk = round(mean(sexBuy$hiv_risk, na.rm = TRUE), 1),
  sd_hiv_risk = round(sd(sexBuy$hiv_risk, na.rm = TRUE), 1),
  avg_total_ace = round(mean(sexBuy$total_ace, na.rm = TRUE), 1),
  sd_total_ace = round(sd(sexBuy$total_ace, na.rm = TRUE), 1)
)

# Calculate stratified averages based on transSex
stratified_stats <- sexBuy %>%
  group_by(transSex) %>%
  summarise(
    avg_alcoholdays = round(mean(alcdays_w_no_drink, na.rm = TRUE), 1),
    sd_alcoholdays = round(sd(alcoholdays, na.rm = TRUE), 1),
    avg_hiv_risk = round(mean(hiv_risk, na.rm = TRUE), 1),
    sd_hiv_risk = round(sd(hiv_risk, na.rm = TRUE), 1),
    avg_total_ace = round(mean(total_ace, na.rm = TRUE), 1),
    sd_total_ace = round(sd(total_ace, na.rm = TRUE), 1)
  )


# Display the table
overall_stats
stratified_stats

```

```{r}
# Assuming 'sexBuy' is your dataset
sexBuy$hfiascat

# Calculate counts and percentages for each variable
table_data <- sexBuy %>%
  dplyr::select(age_below_30, sex_ever, hivstat, hist_sex_b, hfiascat,current_abuse_b, binge_drink, married, ace_fviol, ace_aphys, transSex) %>%
  gather(key = "variable", value = "value", -transSex) %>%  # Convert to long format
  group_by(variable, transSex) %>%
  summarize(
    count_1 = sum(value == 1, na.rm = TRUE),
    total = n(),
    percent_1 = (count_1 / total) * 100
  ) %>%
  ungroup()

# Create the GT table
gt_table <- table_data %>%
  gt() %>%
  tab_header(
    title = "Counts and Percentages of 1 for Each Variable Stratified by transSex"
  ) %>%
  cols_label(
    variable = "Variable",
    transSex = "transSex",
    count_1 = "Count of 1",
    percent_1 = "Percent of 1 (%)"
  ) %>%
  cols_align(
    align = "center",
    columns = everything()
  ) %>%
  tab_spanner_delim(delim = "_")

# Display the table
gt_table


#Ungrouped
table_data <- sexBuy %>%
  dplyr::select(age_below_30, sex_ever, hivstat, hist_sex_b,hfiascat, current_abuse_b, binge_drink, married, ace_fviol, ace_aphys, transSex) %>%
  gather(key = "variable", value = "value", -transSex) %>%  # Convert to long format
  group_by(variable) %>%
  summarize(
    count_1 = sum(value == 1, na.rm = TRUE),
    total = n(),
    percent_1 = (count_1 / total) * 100
  ) %>%
  ungroup()

# Create the GT table
gt_table <- table_data %>%
  gt() %>%
  tab_header(
    title = "Counts and Percentages of 1 for Each Variable Stratified by transSex"
  ) %>%
  cols_label(
    variable = "Variable",
    count_1 = "Count of 1",
    percent_1 = "Percent of 1 (%)"
  ) %>%
  cols_align(
    align = "center",
    columns = everything()
  ) 

# Display the table
gt_table

```
##Create Variable of Transaction and Commercial Sex
```{r}
table(sexBuy$transSex)

#Rate of Transactional Sex is 9.4% of Data Set 
66/(632+66)

sexBuy$weighted_transSex <- ifelse(sexBuy$pastyr_sexworker == 1 & sexBuy$pastyr_gavegiftsex == 1, 3,
                                   ifelse(sexBuy$pastyr_sexworker == 1, 2,
                                   ifelse(sexBuy$pastyr_gavegiftsex == 1, 1, 0)))

table(sexBuy$weighted_transSex)
```


##Sexual Risk and Transactional Sex
```{r}
library(stats)

#Condom Use
sexBuy$transSex_nocondom <- ifelse(sexBuy$pastyr_gavegiftsex_nocondom == 1 | sexBuy$pastyr_sexworker_nocondom == 1, 1, 0)

table(sexBuy$transSex_nocondom)
41/66

# Fit the models
mod1 <- glm(sexBuy$hivstat ~ sexBuy$transSex, family = binomial())
mod2 <- lm(sexBuy$hiv_risk ~ sexBuy$transSex)
mod3 <- glm(sexBuy$pastyr_nocondom ~ sexBuy$transSex, family = binomial())
mod4 <- glm(sexBuy$circumcised ~ sexBuy$transSex, family = binomial())
mod5 <- lm(sexBuy$sexpartners_pastyr ~ sexBuy$transSex)
mod6 <- lm(sexBuy$sexpartners_lifetime ~ sexBuy$transSex)
mod7 <- glm(sexBuy$pastyr_talkcondom ~ sexBuy$transSex, family = binomial())
mod8 <- glm(sexBuy$pastyr_sti ~ sexBuy$transSex, family = binomial())

# Extract p-values and coefficients from the model summaries
coefficients <- c(
  summary(mod1)$coefficients[2, 1],  # for mod1 (logistic regression) - coefficient
  summary(mod2)$coefficients[2, 1],  # for mod2 (linear regression) - coefficient
  summary(mod3)$coefficients[2, 1],  # for mod3 (logistic regression) - coefficient
  summary(mod4)$coefficients[2, 1],  # for mod4 (logistic regression) - coefficient
  summary(mod5)$coefficients[2, 1],  # for mod5 (linear regression) - coefficient
  summary(mod6)$coefficients[2, 1],  # for mod6 (linear regression) - coefficient
  summary(mod7)$coefficients[2, 1],   # for mod7 (logistic regression) - coefficient
  summary(mod8)$coefficients[2, 1]   # for mod8 (logistic regression) - coefficient
)

p_values <- c(
  summary(mod1)$coefficients[2, 4],  # for mod1 (logistic regression) - p-value
  summary(mod2)$coefficients[2, 4],  # for mod2 (linear regression) - p-value
  summary(mod3)$coefficients[2, 4],  # for mod3 (logistic regression) - p-value
  summary(mod4)$coefficients[2, 4],  # for mod4 (logistic regression) - p-value
  summary(mod5)$coefficients[2, 4],  # for mod5 (linear regression) - p-value
  summary(mod6)$coefficients[2, 4],  # for mod6 (linear regression) - p-value
  summary(mod7)$coefficients[2, 4],
  summary(mod8)$coefficients[2, 1]# for mod7 (logistic regression) - p-value
)

# Apply Bonferroni correction
p_adjusted <- p.adjust(p_values, method = "bonferroni")

# Create the results table with coefficients
results_table <- data.frame(
  Test = c("mod1 (hivstat)", "mod2 (hiv_risk)", "mod3 (pastyr_nocondom)",
           "mod4 (circumcised)", "mod5 (sexpartners_pastyr)", 
           "mod6 (sexpartners_lifetime)", "mod7 (pastyr_talkcondom)", "mod8 (sti)"),
  Coefficient = coefficients,
  P_Value = p_values,
  Adjusted_P_Value = p_adjusted
)

# Print the table
print(results_table)


```









#Part Two - Multivariate logit model

##Exploratory LASSO Modeling
```{r}
library(tidyverse)
library(haven)
library(glmnet)
names(sexBuy)
table(sexBuy$phys_viol_b)

# Remove rows with missing values in the relevant columns
sexBuy_clean <- sexBuy %>%
  dplyr::select(transSex, binge_drink, current_abuse_b, alcoholdays, ace_fviol, hist_sex_b, current_abuse_b, hfiascat, age,  phys_viol_b, mstatus, ace_aemot, newAbuse, total_ace) %>%
  na.omit()

sexBuy_clean$mstatus <- as.factor(sexBuy_clean$mstatus)
# Prepare data
X <- model.matrix(transSex ~ binge_drink  + current_abuse_b + ace_aemot + newAbuse + total_ace + alcoholdays + hist_sex_b +phys_viol_b + current_abuse_b +  hfiascat + age + as.factor(mstatus) , data = sexBuy_clean)[,-1]
y <- sexBuy_clean$transSex

# Perform cross-validated lasso regression
set.seed(123)
cv.lasso <- cv.glmnet(X, y, alpha = 1, family = "binomial")

# Print optimal lambda
print(cv.lasso$lambda.min)

# Fit final lasso model with optimal lambda
lasso_model <- glmnet(X, y, alpha = 1, lambda = cv.lasso$lambda.min, family = "binomial")

# Extract coefficients
lasso_coefficients <- coef(lasso_model)
print(lasso_coefficients)

```

#Final Model 
```{r}
library(sandwich)
library(lmtest)
library(stargazer)
library(car)

sexBuy$mstatus2 <- ifelse(sexBuy$mstatus ==3,0, sexBuy$mstatus)

table(sexBuy$mstatus2)

# Fit the logistic regression model
model <- glm(transSex ~  binge_drink + newAbuse + current_abuse + historical_sex_abuse + phys_viol_b + total_ace + factor(mstatus2) , 
             family = binomial, data = sexBuy)

# Calculate robust standard errors
robust_se <- coeftest(model, vcov = vcovHC(model, type = "HC0"))

# Exponentiate the coefficients to get adjusted odds ratios
exp_coef <- exp(coef(model))

# Create a data frame for adjusted odds ratios and robust standard errors
adjusted_odds_ratios <- data.frame(
  Coefficient = names(exp_coef),
  Adjusted_Odds_Ratio = exp_coef,
  Robust_SE = robust_se[, "Std. Error"]
)

# Print the adjusted odds ratios
print(adjusted_odds_ratios)

# Create a stargazer table with the model
getwd()

setwd("/Users/sarahsowell/Downloads")


stargazer(model, type = "html", out = "pdf2.html",
          coef = list(exp_coef),
  dep.var.labels = c("Engagement in Trans. Sex"),
  covariate.labels = c("Intercept", "Binge Drinking", "Exp. of Phys/Verbal Abuse", "Current Sex Abuse", "Hist. Sex Abuse",  "Current Phys. Violence","Total Ace" , "Married", "Divorced/Widowed"),
          se = list(robust_se[, "Std. Error"]),
          p = list(robust_se[, "Pr(>|z|)"]),
          intercept.bottom = FALSE,
          notes = "Robust standard errors in parentheses. Coefficients are exponentiated (Adjusted Odds Ratios).")


```

#Testing for Mediation - Results not Signficant
```{r}
library(bda)
table(sexBuy$total_ace)

sexBuy$high_ace <- ifelse(sexBuy$total_ace > 3,1,0)
# Mediator model (historical_sex_abuse -> binge_drink)
med_model <- glm(binge_drink ~ high_ace,
                 family = binomial, data = sexBuy)

# Outcome model (historical_sex_abuse + binge_drink -> transSex)
out_model <- glm(transSex ~ high_ace + binge_drink,
                 family = binomial, data = sexBuy)

# Extract the coefficients and standard errors
a <- coef(med_model)["high_ace"]
b <- coef(out_model)["binge_drink"]
sa <- summary(med_model)$coefficients["high_ace", "Std. Error"]
sb <- summary(out_model)$coefficients["binge_drink", "Std. Error"]

# Perform Sobel test
sobel_test <- function(a, b, sa, sb) {
  sobel_statistic <- (a * b) / sqrt(b^2 * sa^2 + a^2 * sb^2)
  p_value <- 2 * (1 - pnorm(abs(sobel_statistic)))
  return(list(statistic = sobel_statistic, p_value = p_value))
}

# Calculate Sobel test statistic and p-value
sobel_result <- sobel_test(a, b, sa, sb)

# Print the results
print(paste("Sobel test statistic:", sobel_result$statistic))
print(paste("P-value:", sobel_result$p_value))


model <- glm(transSex ~ binge_drink*total_ace, data = sexBuy)


```

```{r}
# Load necessary packages
library(ggplot2)
library(dplyr)
library(effects)

# Fit the logistic regression model
model <- glm(transSex ~ binge_drink * total_ace, data = sexBuy, family = binomial)

# Create a new data frame for predictions
new_data <- expand.grid(
  binge_drink = unique(sexBuy$binge_drink),
  total_ace = seq(min(sexBuy$total_ace), max(sexBuy$total_ace), length.out = 100)
)

# Generate predicted probabilities
new_data$predicted_prob <- predict(model, newdata = new_data, type = "response")

# Plot the interaction
ggplot(new_data, aes(x = total_ace, y = predicted_prob, color = factor(binge_drink))) +
  geom_line(size = 1) +
  labs(
    title = "Interaction Effect of Binge Drinking and Total ACE on Transactional Sex",
    x = "Total ACE",
    y = "Predicted Probability of Transactional Sex",
    color = "Binge Drinking"
  ) +
  theme_minimal()

```



#Examining Multicolinearity 
```{r}
library(car)

# VIF - Background Factors & Alcohol
linear_model_revised <- (lm(transSex ~ binge_drink  +  ace_fviol + historical_sex_abuse + ace_aphys + total_ace + current_abuse + age, data = sexBuy))



vif_values_revised <- vif(linear_model_revised)
print(vif_values_revised)
```




#Rates of Sexual Abuse
```{r}
sexBuy$rates_abuse <- ifelse(sexBuy$current_abuse_b == 1 & sexBuy$hist_sex_b == 1, 3,
                       ifelse(sexBuy$current_abuse_b == 1, 2,
                              ifelse(sexBuy$hist_sex_b == 1, 1, 0)))

table(sexBuy$rates_abuse)
8/694

(54+50+8)/694

```




